---
title: "modelStudio - perks and features"
author: "Hubert Baniecki"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{modelStudio - perks and features}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning = FALSE,
  message = FALSE
)
```

## More Examples

Let's use `DALEX::HR` dataset to explore some of `modelStudio` parameters.

```{r results="hide"}
train <- DALEX::HR
train$fired <- ifelse(train$status == "fired", 1, 0)
train <- train[, -6]

head(train)
```

```{r echo = FALSE}
knitr::kable(head(train))
```

Prepare data and model for the explainer.

```{r results="hide"}
model <- randomForest::randomForest(fired ~., data = train)

test <- DALEX::HR_test
test$fired <- ifelse(test$status == "fired", 1, 0)
test <- test[, -6]

explainer <- DALEX::explain(model = model,
                            data = test,
                            y = test[, 6])

library("dime")
```

### local explanations

You can pass data points to `new_observation` parameter for local explanations
such as Break Down, Shapley Values and Ceteris Paribus Profiles.

```{r eval = FALSE}
new_observations <- test[1:3,]
rownames(new_observations) <- c("John Snow","Arya Stark","Samwell Tarly")
modelStudio(explainer, new_observation = new_observations)
```

### modelStudio size

You can achieve bigger or smaller `modelStudio` grid with `facet_dim` parameter.

```{r eval = FALSE}
modelStudio(explainer, new_observation = test[1:2,], facet_dim = c(1,2))

modelStudio(explainer, new_observation = test[1:2,], facet_dim = c(3,3))
```

### modelStudio looks

You can manipulate `time` parameter to set animation length. Value 0 will make 
them invisible.

```{r eval = FALSE}
modelStudio(explainer, new_observation = test[1:2,], time = 1000)

modelStudio(explainer, new_observation = test[1:2,], time = 0)
```

### more calculations means more time

You can decrease `N` and `B` parameters to lower computation time or increase
them to get more accurate empirical results.

```{r eval = FALSE}
modelStudio(explainer, new_observation = test[1:2,], N = 100, B = 10)

modelStudio(explainer, new_observation = test[1:2,], N = 1000, B = 50)
```

### progress bar

You can hide computation progress bar messages with `show_info` parameter.

```{r eval = FALSE}
modelStudio(explainer, new_observation = test[1:2,], show_info = FALSE)
```

## Parallel Computation

You can speed up `modelStudio` computation with `parallel` parameter.

```{r}
#TBD
```

## Python scikit-learn Model

You can use `DALEXtra::explain_*()` functions to explain various models.
Bellow basic example of making `modelStudio` for scikit-learn model using
[`DALEXTRA::explain_scikitlearn()`](https://modeloriented.github.io/DALEXtra/reference/explain_scikitlearn.html)

```{r}
library(DALEXtra)
#TBD
```

`modelStudio` should work with any `explainer` class object. Find more about 
making those [here](https://github.com/ModelOriented/DALEXtra)


## Plot Options

TBD
