---
title: "modelStudio - perks and features"
author: "Hubert Baniecki"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{modelStudio - perks and features}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning = FALSE,
  message = FALSE
)
```

## modelStudio parameters

Let's use `DALEX::HR` dataset to explore `modelStudio` parameters:

```{r results="hide"}
train <- DALEX::HR[1:100,]
train$fired <- ifelse(train$status == "fired", 1, 0)
train <- train[,-6]

head(train)
```

```{r echo = FALSE}
knitr::kable(head(train))
```

Prepare data and model for the explainer:

```{r results="hide"}
library("randomForest")
model <- randomForest(fired ~., data = train)

test <- DALEX::HR_test[1:100,]
test$fired <- ifelse(test$status == "fired", 1, 0)
test <- test[,-6]

explainer <- DALEX::explain(model = model,
                            data = test,
                            y = test[,6],
                            verbose = FALSE)

library("dime")
```

### local explanations

You can pass data points to `new_observation` parameter for local explanations
such as Break Down, Shapley Values and Ceteris Paribus Profiles.

```{r eval = FALSE}
new_observations <- test[1:3,]
rownames(new_observations) <- c("John Snow","Arya Stark","Samwell Tarly")
modelStudio(explainer, new_observation = new_observations)
```

### grid size

You can achieve bigger or smaller `modelStudio` grid with `facet_dim` parameter.

```{r eval = FALSE}
modelStudio(explainer, new_observation = test[1:2,], facet_dim = c(1,2))

modelStudio(explainer, new_observation = test[1:2,], facet_dim = c(3,3))
```

### animations

You can manipulate `time` parameter to set animation length. Value 0 will make 
them invisible.

```{r eval = FALSE}
modelStudio(explainer, new_observation = test[1:2,], time = 1000)

modelStudio(explainer, new_observation = test[1:2,], time = 0)
```

### more calculations means more time

You can decrease `N` and `B` parameters to lower computation time or increase
them to get more accurate empirical results.

```{r eval = FALSE}
modelStudio(explainer, new_observation = test[1:2,], N = 100, B = 10)

modelStudio(explainer, new_observation = test[1:2,], N = 1000, B = 50)
```

### progress bar

You can hide computation progress bar messages with `show_info` parameter.

```{r eval = FALSE}
modelStudio(explainer, new_observation = test[1:2,], show_info = FALSE)
```

-------------------------------------------------------------------

## Parallel Computation

You can speed up `modelStudio` computation by setting `parallel` parameter to `TRUE`. 
It uses [`parallelMap`](https://www.rdocumentation.org/packages/parallelMap) package
to calculate local explainers faster. It is really useful when using `modelStudio` with
complicated models, vast datasets or simply many observations are beeing processed. 

All options can be set outside of function call.
More on that [here](https://github.com/berndbischl/parallelMap#being-lazy-configuration).
```{r eval = FALSE}
options(
  parallelMap.default.mode        = "socket",
  parallelMap.default.cpus        = 4,
  parallelMap.default.show.info   = FALSE
)

modelStudio(explainer, new_observation = test[1,], parallel = TRUE)
```

--------------------------------------------------------------------

## Plot Options

You can customize some of `modelStudio` looks by overwriting default options returned
by `getOptions()`.

```{r eval = FALSE}
new_options <- getOptions(
  show_subtitle = TRUE,
  bd_subtitle = "Hello World",
  line_size = 5,
  point_size = 9,
  line_color = "pink",
  point_color = "purple",
  bd_positive_color = "yellow",
  bd_negative_color = "orange" 
)

modelStudio(explainer, new_observation = test[1,],
            options = new_options)
```

-------------------------------------------------------------------

## Python scikit-learn Model

You can use `DALEXtra::explain_*()` functions to explain various models.
Bellow basic example of making `modelStudio` for scikit-learn model using `DALEXtra::explain_scikitlearn()`.

```{r eval = FALSE}
library(DALEXtra)
#TBD
```

`modelStudio` should work with any `explainer` class object. Find more about 
making those [here](https://modeloriented.github.io/DALEXtra/).

