---
title: "modelStudio - perks and features"
author: "Hubert Baniecki"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{modelStudio - perks and features}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning = FALSE,
  message = FALSE
)
```

`dime::modelStudio` uses your model, data and new observations, to provide local
and global explanations. It generates plots and descriptions in the form
of the serverless HTML site, that supports animations and interactivity made with D3.js.

Let's use `DALEX::HR` dataset to explore `modelStudio` parameters:

```{r results="hide"}
train <- DALEX::HR[1:100,]
train$fired <- ifelse(train$status == "fired", 1, 0)
train <- train[,-6]

head(train)
```

```{r echo = FALSE}
knitr::kable(head(train))
```

Prepare data and model for the explainer:

```{r results="hide", eval = FALSE}
# create a random forest model
library("randomForest")
model <- randomForest(fired ~., data = train)

# prepare validation dataset
test <- DALEX::HR_test[1:100,]
test$fired <- ifelse(test$status == "fired", 1, 0)
test <- test[,-6]

# create an explainer
explainer <- DALEX::explain(model = model,
                            data = test[,-6],
                            y = test[,6],
                            verbose = FALSE)

# start dime
library("dime")
```

-------------------------------------------------------------------

## modelStudio parameters

### local explanations

You can pass data points to `new_observation` parameter for local explanations
such as Break Down, Shapley Values and Ceteris Paribus Profiles.

```{r eval = FALSE}
new_observations <- test[1:3,]
rownames(new_observations) <- c("John Snow", "Arya Stark", "Samwell Tarly")

modelStudio(explainer, new_observation = new_observations)
```

### grid size

You can achieve bigger or smaller `modelStudio` grid with `facet_dim` parameter.

```{r eval = FALSE}
# small dashboard with 2 panels
modelStudio(explainer, new_observation = test[1:2,], facet_dim = c(1,2))

# large dashboard with 9 panels
modelStudio(explainer, new_observation = test[1:2,], facet_dim = c(3,3))
```

### animations

You can manipulate `time` parameter to set animation length. Value 0 will make 
them invisible.

```{r eval = FALSE}
# turn on animations
modelStudio(explainer, new_observation = test[1:2,], time = 1000)

modelStudio(explainer, new_observation = test[1:2,], time = 0)
```

### more calculations means more time

`N` is a number of observations used for calculation of partial dependency profiles.
`B` is a number of random paths used for calculation of shapley values.
You can decrease `N` and `B` parameters to lower computation time or increase
them to get more accurate empirical results.

```{r eval = FALSE}
modelStudio(explainer, new_observation = test[1:2,], N = 100, B = 10)

modelStudio(explainer, new_observation = test[1:2,], N = 1000, B = 50)
```

### progress bar

You can hide computation progress bar messages with `show_info` parameter.

```{r eval = FALSE}
modelStudio(explainer, new_observation = test[1:2,], show_info = FALSE)
```

### viewer or browser?

Best described here: [`r2d3 viewer argument`](https://rstudio.github.io/r2d3/articles/visualization_options.html#viewer).

```{r eval = FALSE}
modelStudio(explainer, new_observation = test[1:2,], viewer = "browser")
```

-------------------------------------------------------------------

## parallel computation

You can speed up `modelStudio` computation by setting `parallel` parameter to `TRUE`. 
It uses [`parallelMap`](https://www.rdocumentation.org/packages/parallelMap) package
to calculate local explainers faster. It is really useful when using `modelStudio` with
complicated models, vast datasets or simply many observations are beeing processed. 

All options can be set outside of function call.
More on that [here](https://github.com/berndbischl/parallelMap#being-lazy-configuration).

```{r eval = FALSE}
#set up the cluster 
options(
  parallelMap.default.mode        = "socket",
  parallelMap.default.cpus        = 4,
  parallelMap.default.show.info   = FALSE
)

# calculations will be distributed into 4 cores
modelStudio(explainer, new_observation = test[1,], parallel = TRUE)
```

--------------------------------------------------------------------

## plot options

You can customize some of `modelStudio` looks by overwriting default options returned
by `modelStudioOptions()`.

```{r eval = FALSE}
# set additional graphical parameters
new_options <- modelStudioOptions(
  show_subtitle = TRUE,
  bd_subtitle = "Hello World",
  line_size = 5,
  point_size = 9,
  line_color = "pink",
  point_color = "purple",
  bd_positive_color = "yellow",
  bd_negative_color = "orange" 
)

modelStudio(explainer, new_observation = test[1,],
            options = new_options)
```

-------------------------------------------------------------------

## R mlr model 

You can use `DALEXtra::explain_*()` functions to explain various models.
Bellow basic example of making `modelStudio` for mlr model using `DALEXtra::explain_mlr()`

```{r eval = FALSE}
library(DALEXtra)
# prepare mlr model
task <- mlr::makeRegrTask(id = "task",
                          data = train,
                          target = "fired")

learner <- mlr::makeLearner("regr.randomForest",
                            par.vals = list(ntree = 300),
                            predict.type = "response")

model <- mlr::train(learner, task)

# create an explainer for mlr model
explainer_mlr <- explain_mlr(model, data = test[,-6], y = test[,6])

# call model studio for mlr model
modelStudio(explainer_mlr,
            new_observation = test[1:4,],
            N = 100, B = 10)
```

-------------------------------------------------------------------

## Python scikit-learn model

Bellow basic example of making `modelStudio` for scikit-learn model using `DALEXtra::explain_scikitlearn()`
and **Python Anaconda**.

```{r eval = FALSE}
library(DALEXtra)

titanic_test <- read.csv(system.file("extdata", "titanic_test.csv", package = "DALEXtra")) 
titanic_train <- read.csv(system.file("extdata", "titanic_train.csv", package = "DALEXtra"))

# read scikitlearn model
yml <- system.file("extdata", "scikitlearn.yml", package = "DALEXtra")
pkl_gbm <- system.file("extdata", "scikitlearn.pkl", package = "DALEXtra")
pkl_SGDC <- "SGDC.pkl"

# prepare an explainer for scikitlearn model
explainer_scikit <- explain_scikitlearn(pkl_gbm,
                                        yml = yml,
                                        data = titanic_test[,1:17],
                                        y = titanic_test[,18])

# start model studio
modelStudio(explainer_scikit,
            new_observation = titanic_test[1:4,1:17],
            N = 100, B = 10, options = modelStudioOptions(margin_left = 160))
```

`modelStudio` should work with any `explainer` class object. Find more about 
making those [here](https://modeloriented.github.io/DALEXtra/reference/).
